---
title: " Association Rules"
author: "Joe Kirika"
date: "15/07/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Importing And Previewing The Data          
### Importing the data

```{r}
# importing the library
library(arules)
```

```{r}
# reading the dataset

df2 <- read.transactions('http://bit.ly/SupermarketDatasetII',sep = ",")
```

### Previewing the data          
```{r}
#checking the dataset

print(df2)
```

```{r}
# Verifying the object's class

class(df2)
```

```{r}
# Previewing our first 5 transactions
#
inspect(df2[1:5])
```

```{r}
# To preview the items that make up our dataset,
# 
items<-as.data.frame(itemLabels(df2))
colnames(items) <- "Item"

# the first 10 items
head(items, 10)  
```
```{r}
# Generating a summary of the dataset

summary(df2)
```

```{r}
# Exploring the frequency of some articles (the first ten itemson the list)
# note: they are listred alphabetically
itemFrequency(df2[, 1:10],type = "absolute")
round(itemFrequency(df2[, 1:10],type = "relative")*100,2)
```

```{r}
# Displaying top 10 most common items in the transactions dataset 
# and the items whose relative importance is at least 10%
# 
par(mfrow = c(1, 2))

# plot the frequency of items
itemFrequencyPlot(df2, topN = 10,col="lightgreen")

itemFrequencyPlot(df2, support = 0.1,col="lightblue")

```
  
> the graph on the left shows top common items in the transactions dataset whereas the graph on the left showsitems whose relative importance is at least 10% with mineral water as the lead  

```{r}
# Building a model based on association rules 
# using the apriori function 
# We use Min Support as 0.001 and confidence as 0.8
# ---
# 
rules <- apriori (df2, parameter = list(supp = 0.001, conf = 0.8))
rules
```
> we get a set of 74 rules using a support of 0.001 and confidence of 0.8         

```{r}
# We can perform an exploration of our model 
# through the use of the summary function
summary(rules)
```

> we have some 42 rules with 4 items, 16 rules with 5 items, 15 rules with 3 items and finally 1 rule with 6 items              

```{r}
# Ordering these rules by level of lift
# then looking at the first ten rules.
# ---
# 
rules<-sort(rules, by="lift", decreasing=TRUE)

inspect(rules[1:10])
```


> sorting the rules by lift and two noticeable rules with confidence > 95%         
* {french fries, mushroom cream sauce,pasta}   => {escalope}            
* {mushroom cream sauce, pasta}                => {escalope}
*  {cake, meatballs,mineral water}             => {milk}

### conclusions and Recommendations           
> most common item in the transactions dataset is mineral water   
> we got a set of 74 rules using a support of 0.001 and confidence of 0.8                  
> we have some 42 rules with 4 items, 16 rules with 5 items, 15 rules with 3 items and finally 1 rule with 6 items          
